---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = FALSE, prompt = FALSE, progress = FALSE)

suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(bigrquery))
suppressPackageStartupMessages(library(DBI))
suppressPackageStartupMessages(library(trelliscopejs))
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(maps))
suppressPackageStartupMessages(library(rnaturalearth))
suppressPackageStartupMessages(library(ggsci))
library(ggsci)

BQ_connection <-  dbConnect(dbi_driver(), dataset = "", project = "world-fishing-827")

AMP_v5 <- st_read("AMPS_shapefiles/AMPs_5_v1.shp")
```

```{r message = FALSE}
plotly::ggplotly(ggplot()+
  geom_sf(data = AMP_v5, aes(key =OBJECTID_1) ))

id_lookup <- AMP_v5 

st_geometry(id_lookup) <- NULL

id_lookup %>% 
  select(id = OBJECTID_1, description = Descrip, description_2 = Descripc, area_km2 = AreaKM2) %>% 
  write_csv("mpa_id_lookup.csv")
```

# All effort by all Argentinian vessels

First, lets summarize all fishign effort by argentinian vessels. This will serve as the reference point to compare effort in the proposed MPAs.

```{sql, connection = BQ_connection, output.var = "summary_all_effort_by_ARG_vessel", eval = F}
SELECT
  a.year year,
  a.mmsi mmsi,
  b.gear_type gear_type,
  SUM(hours) total_hours,
  SUM(IF(nnet_score == 1, hours, 0 )) total_fishing_hours,
  SUM(IF(eez_name == "Argentina" AND a.nnet_score == 1, a.hours, 0 )) fishing_hours_eez,
  SUM(IF(eez_name == "Falkland Islands" AND a.nnet_score == 1, a.hours, 0 )) fishing_hours_falkland,
  SUM(IF(eez_name IS NULL AND (distance_from_shore >= 10*1852 OR FAO_region IN (88, 48,58)) AND a.nnet_score == 1, a.hours, 0 )) fishing_hours_hs,
FROM (
  SELECT
    *,
    year(timestamp) year,
     Integer(REGEXP_REPLACE( if(REGEXP_EXTRACT(regions,'\"(fao:.*?)\"') contains ".",  LEFT(REGEXP_EXTRACT(regions,'\"(fao:.*?)\"'),INSTR(REGEXP_EXTRACT(regions,'\"(fao:.*?)\"'),".")-1),REGEXP_EXTRACT(regions,'\"(fao:.*?)\"')), '[^0-9 ]','')) FAO_region,
  FROM
    [world-fishing-827:gfw_research.nn]
  WHERE
    _PARTITIONTIME BETWEEN TIMESTAMP('2014-01-01')
    AND TIMESTAMP('2016-12-31')
    AND lat < 90
    AND lat > -90
    AND lon < 180
    AND lon >- 180
    AND seg_id IN (SELECT seg_id FROM [world-fishing-827:gfw_research.good_segments]) ) a
INNER JOIN (
  SELECT
    year,
    mmsi,
    flag_country_name,
    flag_iso3,
    sovereign_flag_country_name,
    sovereign_flag_iso3,
    gear_type_all_years gear_type,
    length,
    tonnage,
    engine_power,
  FROM
    [high-seas:vessel_characteristics.VC_all_years]
  WHERE
    flag_country_name == "Argentina"
  GROUP BY
    year,
    mmsi,
    flag_country_name,
    flag_iso3,
    sovereign_flag_country_name,
    sovereign_flag_iso3,
    length,
    tonnage,
    engine_power,
    gear_type)b
ON
  a.mmsi = b.mmsi
  AND a.year = b.year
GROUP BY
  year,
  mmsi,
  gear_type
  having total_fishing_hours > 0
```

### All fishing positions within the EEZ

Now lets get all **fishing** positions by these vessels inside the EEZ. We will then clip them to the MPAS.

Since the resulting files are huge, we need to run a query for each year separately: 

```{sql connection = BQ_connection, output.var = "all_fishing_positions_in_eez_2016", eval = F}
SELECT
year(timestamp) year,
mmsi, 
lat, 
lon,
hours
  FROM
    [world-fishing-827:gfw_research.nn]
  WHERE
    _PARTITIONTIME BETWEEN TIMESTAMP('2016-01-01')
    AND TIMESTAMP('2016-12-31')
    AND lat < 90
    AND lat > -90
    AND lon < 180
    AND lon >- 180
    AND nnet_score == 1
    AND (eez_name == "Argentina" or eez_name == "Falkland Islands")
    AND seg_id IN (SELECT seg_id FROM [world-fishing-827:gfw_research.good_segments]) 
    AND mmsi IN (Select mmsi FROM [high-seas:vessel_characteristics.VC_all_years] WHERE flag_country_name == "Argentina")
```

```{sql connection = BQ_connection, output.var = "all_fishing_positions_in_eez_2015", eval = F}
SELECT
year(timestamp) year,
mmsi, 
lat, 
lon,
hours
  FROM
    [world-fishing-827:gfw_research.nn]
  WHERE
    _PARTITIONTIME BETWEEN TIMESTAMP('2015-01-01')
    AND TIMESTAMP('2015-12-31')
    AND lat < 90
    AND lat > -90
    AND lon < 180
    AND lon >- 180
    AND nnet_score == 1
    AND (eez_name == "Argentina" or eez_name == "Falkland Islands")
    AND seg_id IN (SELECT seg_id FROM [world-fishing-827:gfw_research.good_segments]) 
    AND mmsi IN (Select mmsi FROM [high-seas:vessel_characteristics.VC_all_years] WHERE flag_country_name == "Argentina")
```

```{sql connection = BQ_connection, output.var = "all_fishing_positions_in_eez_2014", eval = F}
SELECT
year(timestamp) year,
mmsi, 
lat, 
lon,
hours
  FROM
    [world-fishing-827:gfw_research.nn]
  WHERE
    _PARTITIONTIME BETWEEN TIMESTAMP('2014-01-01')
    AND TIMESTAMP('2014-12-31')
    AND lat < 90
    AND lat > -90
    AND lon < 180
    AND lon >- 180
    AND nnet_score == 1
    AND (eez_name == "Argentina" or eez_name == "Falkland Islands")
    AND seg_id IN (SELECT seg_id FROM [world-fishing-827:gfw_research.good_segments]) 
    AND mmsi IN (Select mmsi FROM [high-seas:vessel_characteristics.VC_all_years] WHERE flag_country_name == "Argentina")
```


```{r eval = F}
all_fishing_positions_in_eez <- bind_rows(all_fishing_positions_in_eez_2014, all_fishing_positions_in_eez_2015, all_fishing_positions_in_eez_2016)

write_csv(all_fishing_positions_in_eez, "saved_files/all_fishing_positions_in_eez.csv")

write_csv(summary_all_effort_by_ARG_vessel, "saved_files/summary_all_effort_by_ARG_vessel.csv")
```

Now lets do point in polygon operations to assign to each fishing position the MPA in which it occurs (if any)


```{r, eval = F}
all_fishing_positions_in_eez <- read_csv("saved_files/all_fishing_positions_in_eez.csv")

all_fishing_positions_in_eez_SF = st_as_sf(all_fishing_positions_in_eez, coords = c("lon", "lat"), 
                 crs = st_crs(AMP_v5))

all_positions_within_MPAs <- st_join(all_fishing_positions_in_eez_SF, AMP_v5) %>% 
  filter(!is.na(OBJECTID_1))

all_positions_within_MPAs_df <- all_positions_within_MPAs
st_geometry(all_positions_within_MPAs_df) <- NULL
```

```{r eval = F}
summary_of_effort_within_MPAs <- all_positions_within_MPAs_df %>% 
  group_by(year, mmsi, OBJECTID_1) %>% 
  summarise(fishing_hours = sum(hours)) %>% 
  spread(OBJECTID_1, fishing_hours) 

summary_of_effort_within_MPAs[is.na(summary_of_effort_within_MPAs)] <- 0

colnames(summary_of_effort_within_MPAs) <- c("year", "mmsi", "fishing_hours_mpa_1", "fishing_hours_mpa_2", "fishing_hours_mpa_3", "fishing_hours_mpa_4", "fishing_hours_mpa_5", "fishing_hours_mpa_6", "fishing_hours_mpa_7","fishing_hours_mpa_8",  "fishing_hours_mpa_10", "fishing_hours_mpa_11", "fishing_hours_mpa_12", "fishing_hours_mpa_13", "fishing_hours_mpa_14")

write_csv(summary_of_effort_within_MPAs,"saved_files/summary_of_effort_by_mmsi_within_MPAs.csv")
```

### Summary and visualizations

```{r}
summary_all_effort_by_ARG_vessel <- read_csv("saved_files/summary_all_effort_by_ARG_vessel.csv")

summary_of_effort_within_MPAs <- read_csv("saved_files/summary_of_effort_by_mmsi_within_MPAs.csv")

summary_all_effort_by_ARG_vessel <- summary_all_effort_by_ARG_vessel %>% 
  left_join(summary_of_effort_within_MPAs, by = c("year", "mmsi"))

summary_all_effort_by_ARG_vessel[is.na(summary_all_effort_by_ARG_vessel)] <- 0

summary_all_effort_by_ARG_vessel <- summary_all_effort_by_ARG_vessel %>%
  mutate(fishing_hours_all_mpas =  rowSums(.[9:21]))
```

#### How much effort, relative to entire EEZ happens in each MPA?

```{r}
vessel_and_effort_by_mpa <- summary_all_effort_by_ARG_vessel %>% 
  select(-fishing_hours_eez, -total_hours,-fishing_hours_falkland, -fishing_hours_hs) %>% 
  gather(region, fishing_hours, - year, -mmsi, -gear_type, -total_fishing_hours) %>% 
  ungroup() %>% 
  mutate(MPA_id = as.integer(stringr::str_extract(region,"[0-9]+"))) %>%
  arrange(MPA_id) %>% 
  replace_na(list(MPA_id = "all")) %>% 
  mutate(MPA_id = factor(MPA_id)) %>% 
  group_by(year, MPA_id) %>% 
  summarise(vessels = n_distinct(mmsi[fishing_hours > 0]),
            fishing_hours_in_MPA = sum(fishing_hours),
            total_fishing_hours_by_affected_vessels =  sum(total_fishing_hours[fishing_hours >0]),
            total_fishing_hours_all_vessels =  sum(total_fishing_hours)) %>% 
  mutate_if(is.numeric, round, 2)
```

```{r}
effort_by_mpa_relative_to_all <- vessel_and_effort_by_mpa %>% 
  select(year, MPA_id, fishing_hours_in_MPA) %>% 
  spread(year, fishing_hours_in_MPA) %>% 
  rename(effort_2014 = `2014`, effort_2015 = `2015`, effort_2016 = `2016`) %>% 
  arrange(desc(effort_2016)) %>% 
  left_join(vessel_and_effort_by_mpa %>% 
              mutate(percent_all_effort = round(100*fishing_hours_in_MPA/total_fishing_hours_all_vessels, 2)) %>% 
              select(year, MPA_id, percent_all_effort) %>% 
              spread(year, percent_all_effort) %>% 
              rename(percent_2014 = `2014`, percent_2015 = `2015`, percent_2016 = `2016`)) 

effort_by_mpa_relative_to_all %>% 
transmute(MPA_id = MPA_id,
            `fishing hours in 2014` = paste(effort_2014, " (", percent_2014, "%)", sep = ""),
            `fishing hours in 2015` = paste(effort_2015," (", percent_2015, "%)", sep = ""),
            `fishing hours in 2016` = paste(effort_2016," (", percent_2016, "%)", sep = "")) %>% 
  write_csv("saved_files/table_of_effort_by_mpa_relative_to_all.csv")
```

```{r}
effort_by_mpa_relative_to_all_plot <- effort_by_mpa_relative_to_all %>% 
  select(MPA_id, percent_2016) %>% 
  ggplot(aes(x = fct_reorder(MPA_id,percent_2016) , y = percent_2016, fill = percent_2016))+
  geom_col(show.legend = FALSE)+
  theme_minimal()+
  labs(x = "MPA id", y = "% of fishing effort")+
  scale_y_continuous(limits = c(0,100))+
  scale_fill_material("teal")+
  labs(title = "Percent of effort in each proposed MPA",
       subtitle = "Relative to total effort",
       caption = 2016)

tiff(paste('saved_plots/effort_by_mpa_relative_to_all_plot.tiff',sep = ""), height = 12, width = 17, units = 'cm', 
     compression = "lzw", res = 300)

print(effort_by_mpa_relative_to_all_plot)
  
invisible(dev.off())
```


#### What is the % of effort within each MPA relative to the effort outside the MPA by the vessels potentially affected?

```{r}
effort_by_mpa_relative_to_effort_from_affected_vessels <- vessel_and_effort_by_mpa %>% 
  select(year, MPA_id, fishing_hours_in_MPA) %>% 
  spread(year, fishing_hours_in_MPA) %>% 
  rename(effort_2014 = `2014`, effort_2015 = `2015`, effort_2016 = `2016`) %>% 
  arrange(desc(effort_2016)) %>% 
  left_join(vessel_and_effort_by_mpa %>% 
              mutate(percent_effort_affected_vessels = round(100*fishing_hours_in_MPA/total_fishing_hours_by_affected_vessels, 2)) %>% 
              select(year, MPA_id, percent_effort_affected_vessels) %>% 
              spread(year, percent_effort_affected_vessels) %>% 
              rename(percent_2014 = `2014`, percent_2015 = `2015`, percent_2016 = `2016`)) 

effort_by_mpa_relative_to_effort_from_affected_vessels %>% 
transmute(MPA_id = MPA_id,
            `fishing hours in 2014` = paste(effort_2014, " (", percent_2014, "%)", sep = ""),
            `fishing hours in 2015` = paste(effort_2015," (", percent_2015, "%)", sep = ""),
            `fishing hours in 2016` = paste(effort_2016," (", percent_2016, "%)", sep = "")) %>% 
  write_csv("saved_files/table_of_effort_by_mpa_relative_to_effort_from_affected_vessels_only.csv")
```

```{r}
effort_by_mpa_relative_to_effort_from_affected_vessels_plot <- effort_by_mpa_relative_to_effort_from_affected_vessels %>% 
  select(MPA_id, percent_2016) %>% 
  ggplot(aes(x = fct_reorder(MPA_id,percent_2016) , y = percent_2016, fill = percent_2016))+
  geom_col(show.legend = FALSE)+
  theme_minimal()+
  labs(x = "MPA id", y = "% of fishing effort")+
  scale_y_continuous(limits = c(0,100))+
  scale_fill_material("teal")+
  labs(title = "Percent of effort in each proposed MPA",
       subtitle = "Relative to effort from affected vessels only",
       caption = 2016)

tiff(paste('saved_plots/effort_by_mpa_relative_to_effort_from_affected_vessels_plot.tiff',sep = ""), height = 12, width = 17, units = 'cm', 
     compression = "lzw", res = 300)

print(effort_by_mpa_relative_to_effort_from_affected_vessels_plot)
  
invisible(dev.off())
```


#### How many vessels fish in each MPA?

```{r}
affected_vessels <- vessel_and_effort_by_mpa %>% 
  select(year, MPA_id, vessels) %>% 
  spread(year, vessels) %>% 
  rename(vessels_2014 = `2014`, vessels_2015 = `2015`, vessels_2016 = `2016`) %>% 
  arrange(desc(vessels_2016)) %>% 
  mutate(p_2014 = 100*vessels_2014/n_distinct(summary_all_effort_by_ARG_vessel$mmsi[summary_all_effort_by_ARG_vessel$year == 2014]),
         p_2015 = 100*vessels_2015/n_distinct(summary_all_effort_by_ARG_vessel$mmsi[summary_all_effort_by_ARG_vessel$year == 2015]),
         p_2016 = 100*vessels_2016/n_distinct(summary_all_effort_by_ARG_vessel$mmsi[summary_all_effort_by_ARG_vessel$year == 2016])) %>% 
  mutate_if(is.numeric, round, digits = 1)


affected_vessels %>% 
transmute(MPA_id = MPA_id,
            `fishing vessels in 2014` = paste(vessels_2014, " (", p_2014, "%)", sep = ""),
            `fishing vessels in 2015` = paste(vessels_2015," (", p_2015, "%)", sep = ""),
            `fishing vessels in 2016` = paste(vessels_2016," (", p_2016, "%)", sep = "")) %>% 
  write_csv("saved_files/affected_vessels_by_MPA.csv")
```

```{r}
(affected_vessels_plot <- affected_vessels %>% 
  select(MPA_id, p_2016) %>% 
  ggplot(aes(x = fct_reorder(MPA_id,p_2016) , y = p_2016, fill = p_2016))+
  geom_col(show.legend = FALSE)+
  theme_minimal()+
  labs(x = "MPA id", y = "% of all fishing vessels")+
  scale_y_continuous(limits = c(0,100))+
  scale_fill_material("teal")+
  labs(title = "Percent of fishing vessels operating in each proposed MPA",
       subtitle = "Relative to all Argentinian flagged vessels",
       caption = 2016))

tiff(paste('saved_plots/affected_vessels_plot.tiff',sep = ""), height = 12, width = 17, units = 'cm', 
     compression = "lzw", res = 300)

print(affected_vessels_plot)
  
invisible(dev.off())
```


### Dsitrbutions of % of effort in each MPA


```{r}
percent_effort_in_MPA_by_mmsi <- summary_all_effort_by_ARG_vessel %>% 
  select(-fishing_hours_eez, -total_hours,-fishing_hours_falkland, -fishing_hours_hs) %>% 
  gather(region, fishing_hours, - year, -mmsi, -gear_type, -total_fishing_hours) %>% 
  ungroup() %>% 
  mutate(MPA_id = as.integer(stringr::str_extract(region,"[0-9]+"))) %>%
  arrange(MPA_id) %>% 
  replace_na(list(MPA_id = "all")) %>% 
  mutate(MPA_id = factor(MPA_id)) %>% 
  mutate(percent_effort_in_MPA = round(100*fishing_hours/total_fishing_hours, 2)) 

(boxplot_v2 <- percent_effort_in_MPA_by_mmsi %>% 
  filter(percent_effort_in_MPA > 0, year == 2016) %>% 
  ggplot() +
  geom_boxplot(aes(x  = fct_reorder(MPA_id, percent_effort_in_MPA), y = percent_effort_in_MPA, fill = MPA_id), show.legend = FALSE) +
  theme_minimal() +
  labs(x = "MPA id", y = "% effort", caption = "2016", title = "Percent of effort in each MPA", subtitle = "included here are vessels that operate in at least one MPA")+
  scale_fill_manual(values = rev(ggpubr::get_palette(palette = "npg", 14))))
  

tiff(paste('saved_plots/boxplot_v2.tiff',sep = ""), height = 12, width = 17, units = 'cm', 
     compression = "lzw", res = 300)

print(boxplot_v2)
  
invisible(dev.off())
```


```{r}
(boxplot_by_gear_type_v2 <- percent_effort_in_MPA_by_mmsi %>% 
  ungroup() %>% 
  filter(percent_effort_in_MPA > 0, year == 2016, gear_type %in% c("purse_seines", "drifting_longlines","trawlers","squid_jigger")) %>% 
  select(MPA_id, gear_type, percent_effort_in_MPA) %>% 
  ggplot() +
  geom_boxplot(aes(x  = fct_reorder(MPA_id, percent_effort_in_MPA), y = percent_effort_in_MPA, fill = gear_type), show.legend = FALSE) +
  theme_minimal() +
  labs(x = "MPA id", 
        y = "% effort", 
        caption = "2016",
        title = "Percent of effort in each MPA", 
        subtitle = "included here are vessels that operate in at least one MPA")+
  scale_fill_manual(values = rev(ggpubr::get_palette(palette = "npg", 4)))+
  facet_wrap("gear_type", scales = 'free'))

tiff(paste('saved_plots/boxplot_by_gear_type_v2.tiff',sep = ""), height = 12, width = 17, units = 'cm', 
     compression = "lzw", res = 300)

print(boxplot_by_gear_type_v2)
  
invisible(dev.off())
```


```{r eval = F}
(exploding_bixplot <- explodingboxplotR::exploding_boxplot(
  percent_effort_in_MPA_by_mmsi %>% 
  filter(year == 2016) %>% 
  filter(percent_effort_in_MPA > 0, gear_type %in% c("purse_seines", "drifting_longlines","trawlers","squid_jigger")),
  y = "percent_effort_in_MPA",
  group = "MPA_id",
  color = "MPA_id",
  margin = list(bottom = 50, left = 50, top = 20, right = 20),
  xlab = "MPA id",
  ylab = "% effort in MPA",
  label = "gear_type",
  iqr = 1.5
))

htmlwidgets::saveWidget(exploding_bixplot, "exploding_boxplot.html")
```


